---
title: "demoCode_dbmdev"
author: "Kym D Perry"
date: "13 October 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, message = FALSE, 
                      dpi = 144, fig.width = 12, fig.height = 8)

# clear workspace, load packages

#rm(list = ls())
library(dbmdev)
library(dplyr)
library(magrittr)
library(readr)
library(readxl)
library(lubridate)
library(reshape2)
library(parallel) # for mclapply
library(ggplot2)


```

## Running phenological models for diamondback moth using custom R package dbmdev

This document demonstrates how to run Briere's temperature-based model to predict the development rate of diamondback moth, Plutella xylostella, using custom R package dbmdev.

Package authors: Kym. D. Perry, Stephen M. Pederson, Michael A. Keller
Version date: October 2016
Maintainer: Kym. D Perry kym.perry@adelaide.edu.au

For documentation and help pages, type ?<functionName>
Index on help pages contains the full list of functions

## Run model for single location

```{r Modelling, single location}

# Read in a raw climate data file containing a time series of temperature observations.
df <- read_csv(
  "C:/UserData/Kym/PhD/Data/climateData/halfHourlyStation2015/HM01X_Data_026100_999999998908531_PadthaWay_South.txt") 
head(df) # the raw df

# Format and aggregate into hourly observations
df %<>% dbmdev::formatBOMstdata()
head(df) # the formatted df

# Create a data.frame with the DBM development parameters required for Briere's model
# Note, DBM parameters were generated in R package `nlsreg` using development data from Liu 2002.
devParams <- data.frame(
  egg = c(0.0003592, 1.754, 35.08, 4.347),
  instar_1_2 = c(0.00036, -2.9122, 32.4565, 60.5092),
  instar_3 = c(0.0009688, 0.7893, 32.04, 14.53),
  instar_4 = c(0.000674, -1.22561, 32.04873, 15.27334),
  prePupa = c(0.00181533, 3.963006, 33.04467, 4.317786),
  pupaMF = c(0.000396921, 2.417172, 32.44556, 11.99131))
row.names(devParams) <- c("a", "Tmin", "Tmax", "m")
(devParams <- devParams %>% t)


# Run forward prediction with input parameters using fwdDev()
f3 <- fwdDev(tempObs = df,             # climate data frame  
             devParamsDf = devParams,  # insect development parameters
             startDate = "2015-03-01", # a string in yyyy-mm-dd format
             startStage = "egg",       # a string matching one of row.names(devParams)
             startDev = 0.5,             # a number between 0 (no dev) and 1 (dev completed)
             gens = 3)                 # consecutive generations to model

# Output as a summary of life stages (each generation is stored as a separate list element)
f3_stages <- fwdDev(tempObs = df,     
             devParamsDf = devParams,  
             startDate = "2015-03-01", 
             startStage = "egg",     
             startDev = 0.5,             
             gens = 3,
             output = "stages")        # output as "stages" (default value is "increments")

# Output as a summary of generations
f3_gens <- fwdDev(tempObs = df,     
             devParamsDf = devParams,  
             startDate = "2015-03-01", 
             startStage = "egg",     
             startDev = 0.5,             
             gens = 3,
             output = "generations")   # output as "generations"

# Viewing output ... for `increments` and `stages` output, each generation is stored as separate dfs in a list

f3
f3_stages
f3_gens

f3_stages_df <- bind_rows(f3_stages)  # to combine generations into a single dataframe 
f3_stages_df

View(bind_rows(f3))       # to view object a scrollable spreadsheet (if list, must bind_rows())

# Run reverse predictions with input parameters using revDev()
r10 <- revDev(tempObs = df,
              devParamsDf = devParams,
              startDate = "2015-03-01",
              startStage = "pupaMF",
              startDev = 0.75,
              gens = 10,
              output = "increments")

r10_gens <- revDev(tempObs = df,
              devParamsDf = devParams,
              startDate = "2015-03-01",
              startStage = "pupaMF",
              startDev = 0.75,
              gens = 10,
              output = "generations")

r10_gens

# To check development curves, plot "increments" output

# fwd ...
ggplot(data = bind_rows(f3), aes(x = obs, y = dev)) + 
  geom_point() +
  facet_wrap(~stage)

# rev...
ggplot(data = bind_rows(r10), aes(x = obs, y = abs(dev))) + 
  geom_point() +
  facet_wrap(~stage)




```

## Running models for multiple locations simultaneously

```{r Running multiple models, warning = FALSE}

# Read in a data frame with parameters for each site stored in separate columns
sitesDf <- read_excel("C:/UserData/Kym/PhD/Data/colonisation/2015-data/sitesDf2015.xlsx") %>% 
  head() %>%
  #mutate(startDate = as.character(startDate)) %>% # because the new version of R parses in datetimes!
  mutate(startDev = 0.5) %>% # added dummy value here, but can be specified as a data column before reading in file
  dplyr::select(site, station, startDate, startStage, startDev)

sitesDf

# Read all temperature time series data files into a list and name each df (list element) by th unique `station` ID
# Important note: climate data for each site are extracted from dfList by matching the unique `station` ID.
wd <- "C:/UserData/Kym/PhD/Data/climateData/halfHourlyStation2015/"
dfs <- list.files(path = wd, pattern = ".txt") %>% 
  paste0(wd, .) %>%
  lapply(X = unique(sitesDf$station), FUN = grep, ., value = TRUE) %>% unlist # extract the station IDs
dfList <- mclapply(dfs, read_csv) %>% mclapply(FUN = formatBOMstdata)
names(dfList) <- lapply(X = dfList, FUN = function(x) {unique(x$station)})

# Run reverse predictions for all sites
mloc_r3 <- mlocDev(locsDf = sitesDf,
                   tempObsDfList = dfList,
                   devParamsDf = devParams,
                   timedir = "rev",
                   gens = 3,
                   output = "increments")

# ... output as stages summary
mloc_r3_stages <- mlocDev(locsDf = sitesDf,
                          tempObsDfList = dfList,
                          devParamsDf = devParams,
                          timedir = "rev",
                          gens = 3,
                          output = "stages")

# ... output as generations summary
mloc_r3_gens <- mlocDev(locsDf = sitesDf,
                          tempObsDfList = dfList,
                          devParamsDf = devParams,
                          timedir = "rev",
                          gens = 3,
                          output = "generations")

# Output is stored in a list with a separate output dataframe for each site 
mloc_r3 
mloc_r3_stages
mloc_r3_gens

# To combine output into a single dataframe ...
bind_rows(mloc_r3_gens)


```


## Dealing with missing climate data 

## Interpolating hourly temperatures from daily temperature data

Phenological models commonly require hourly data measurements. Sometimes, hourly data are unavailable or strings of consecutive hours of temperature observations are missing from a time series. In this situation, models can be used to infer daily diurnal temperature curves from daily maximum and minimum temperature to generate estimated hourly observations.  

```{r Interpolating hourly temperatures}

# Read in daily maxmin temperature obseravtions for weather stations of interest.
# GPS coordinates for each station are also required
stationCrds <- read_csv(  "C:/UserData/Kym/PhD/Data/climateData/weatherStationCoords.csv")
wd <- "C:/UserData/Kym/PhD/Data/climateData/dailyMaxMinStation2015/"

mxpath <- list.files(path = wd, pattern = "_dmax.csv") %>%
  paste0(wd, .)
mnpath <- list.files(path = wd, pattern = "_dmin.csv") %>%
  paste0(wd, .)

# dmax files ...
maxDf <- mclapply(mxpath, read_csv) %>%
  lapply(., mutate,
         date = as.Date(paste(Year, Month, Day, sep = "-"))) %>%
  lapply(., dplyr::select,
         station = contains("station"),
         date,
         max = contains("Maximum", ignore.case = FALSE)) %>%
  bind_rows()

# merge dmax and dmin files by location and date ...
dailyDf <- mclapply(mnpath, read_csv) %>%
  lapply(., mutate,
         date = as.Date(paste(Year, Month, Day, sep = "-"))) %>%
  lapply(., dplyr::select, station = contains("station"), date,
         min = contains("Minimum", ignore.case = FALSE)) %>%
  bind_rows() %>%
  merge(maxDf, by = c("station", "date")) %>%
  merge(stationCrds, ., by = "station") %>%
  arrange(station, date)

head(dailyDf)



# First, calculate sunrise and sunset times for each location and date using DBMdevmode2::sunriset()
dailyDf %<>% dbmdev::msunriset()
head(dailyDf)

# Plot daylength ...
ggplot(dailyDf, aes(x = date, y = daylength, colour = stationName)) + geom_line()

# Plot sunrise and sunset times....
# Note the effect of changes to summer time! On the `change` dates, there is an unexpected blip

df2 <- dailyDf %>%
  melt(id.vars = c("date","stationName"),
       measure.vars = c("sunrise", "sunset"),
       variable.name = "direction",
       value.name = "suntimes")
ggplot(df2, aes(x = date, y = hour(suntimes) + minute(suntimes)/60, colour = direction)) + 
  geom_line() + 
  facet_wrap(~stationName)

# Now, interpolate hourly temperature
hourlyDf <- dbmdev::hourlyTemperatures(dailyDf)

# Plot the entire dataset ...
ggplot(hourlyDf, aes(x = datetime, y = obs)) +
  geom_line() + 
  facet_wrap(~station)

# Plot one month of data ...
dateSeq <- seq(as.Date("2014-01-01"), as.Date("2014-01-31"), by = "day")
ggplot(filter(hourlyDf, lubridate::date(datetime) %in% dateSeq), 
              aes(x = datetime, y = obs)) +
  geom_line() + 
  facet_wrap(~station)

# Plot one week of data
dateSeq2 <- seq(as.Date("2014-01-01"), as.Date("2014-01-08"), by = "day")
ggplot(filter(hourlyDf, lubridate::date(datetime) %in% dateSeq2), 
       aes(x = datetime, y = obs)) +
  geom_line() + 
  facet_wrap(~station)

# Compare against real data ...

wd <- "C:/UserData/Kym/PhD/Data/climateData/halfHourlyStation2015/"
dfs <- list.files(path = wd, pattern = "+.txt") %>% 
  paste0(wd, .)
dfList <- mclapply(dfs, read_csv) %>% 
  mclapply(., formatBOMstdata)

ggplot(hourlyDf %>% filter(lubridate::date(datetime) %in% dateSeq2), 
       aes(x = datetime, y = obs)) +
  geom_line(linetype = "dashed", size = 1, colour = "blue") +
  geom_line(data = bind_rows(dfList) %>% 
              filter(lubridate::date(datetime) %in% dateSeq2), 
            aes(x = datetime, y = obs),
            size = 0.75, colour = "orange", alpha = 0.75) +
  facet_wrap(~station)

# END

```



